---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Load libraries

```{r}
library(readxl)
library(tidyverse)
library(data.table)
library(janitor)
```

```{r}
dat <- read_xlsx("data/Typhoid_outbreak_TS_datasheet_SciData_submission_final.xlsx", 
                       sheet = "Typhoid outbreaks summary")
ts <- read_xlsx("data/Typhoid_outbreak_TS_datasheet_SciData_submission_final.xlsx", 
                       sheet = "Time-series data")
```

```{r}
ts <- janitor::clean_names(ts)
names(ts)
# ts |> 
#   ggplot(aes(start_date, no_of_patients))+
#   geom_line()+
#   facet_wrap(~study_id, scales = "free")

ts$unit_time <- as.integer(as.Date(ts$end_date) - as.Date(ts$start_date)) + 1

# ts[!(ts$unit_time %in% c(1,7,30,31)),] -> d
# unique(d$study_id)
ts |> 
  group_by(study_id) |>
  mutate(time_since_outbreak = cumsum(unit_time)) -> ts
  
```

```{r}
dat <- janitor::clean_names(dat)
dat$amr_status <- ifelse(dat$amr_status == "-", NA, dat$amr_status)
```

```{r}
table(dat$who_region)
table(dat$country)
table(dat$attack_rate_percent)
sum(as.numeric(dat$total_suspected_cases), na.rm=T)
sum(as.numeric(dat$total_confirmed_cases), na.rm=T)
```

```{r}
dat$intervention <- ifelse(dat$intervention == "-", NA, dat$intervention)
table(dat$intervention)
```

outbreak size

```{r}
ts |> 
  group_by(study_id) |>
  summarize(outbreak_size = sum(no_of_patients)) -> d

ggplot(d)+
  geom_histogram(aes(x=outbreak_size), 
                 breaks = seq(0, max(d$outbreak_size), by = 100))
```

outbreak duration

```{r}
ts |>
  group_by(study_id) |> 
  summarize(outbreak_size = sum(no_of_patients),
            outbreak_duration = as.integer(max(end_date) - min(start_date))) -> d

ggplot(d)+
  geom_histogram(aes(x=as.integer(outbreak_duration)), 
                 breaks = seq(0, max(d$outbreak_duration), by = 30))
# summarize(num_case = sum(suspected, na.rm=T))

```

outbreak time series by study

```{r}
ts |>
  group_by(study_id) |> 
  mutate(id = cur_group_id()) -> ts
 
ggplot(ts)+
  geom_line(aes(x=time_since_outbreak/7, y=suspected, group=study_id, 
                color=as.factor(id)))
ts |> 
  group_by(study_id) |> 
  summarize(outbreak_duration = as.integer(max(end_date) - min(start_date))) |>
  ggplot(aes(outbreak_duration/7)) +
  stat_ecdf(geom = "step")
```



```{r}
# 340 - Lewis 2005
d <- ts[ts$study_id == "#340 - Lewis 2005",]
```

Exclude the studies for modeling based on the following criteria


Studies 

1. #24 - Al-Sanouri 2008
2. #257 – Imanishi 2014, Dzivaresekwa
3. #27 - Ali 2017
_4. #276 - Kabwama 2017_
5. #293 – Keddy 2011
6. #340 - Lewis 2005
7. #353 - Lutterloh 2012
8. #414 - Muti 2014
9. #417 - N'Cho 2019
10. #458 - Polonsky 2014, Dzivaresekwa
11. #470 - Qamar 2018
12. #52 - Aye 2004
13. #619 – Yousafzai 2019
14. #67 – Bano-Zaidi 2018

1. Kabwama SN, Bulage L, Nsubuga F, Pande G, Oguttu DW, Mafigiri R, et al. A large and persistent outbreak of typhoid fever caused by consuming contaminated water and street-vended beverages: Kampala, Uganda, January – June 2015. BMC Public Health. 2017 Jan 5;17(1):23. 

1. Nimonkar RA, Goyal AK, Ahmed S, Pardal MPS, Singh S. Clinico-epidemiological study of an outbreak of typhoid in North India. J Family Med Prim Care. 2022;11(7):3570–4. 

1. Srinivasan M, Sindhu KN, Kumar JS, Ramasamy RK, Pragasam AK, Aasaithampi P, et al. Outbreak of Typhoid Fever in Children of Urban Vellore: A Report from the Surveillance for Enteric Fever in India Cohort. Am J Trop Med Hyg. 2022;107(1):82–5. 

