---
title: "Vaccine impact"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

### Load packages

```{r}
# devtools::load_all() # works only for me who is writing a package
set.seed(42)
library(Cairo)
library(data.table)
library(tidyverse)
library(readxl)
source("R/functions.R") # utility functions

# figure size
fgw1 <- figure_size[figure_size$journal =="Lancet",]$single
fgw2 <- figure_size[figure_size$journal =="Lancet",]$double
```

### Load outbreak datasets

```{r}
ts <- read_xlsx("data/Typhoid_outbreak_TS_datasheet_SciData_submission_final.xlsx", 
                       sheet = "Time-series data")
# 340 - Lewis 2005
d <- ts[ts$study_id == "#340 - Lewis 2005",]
# extract ids for the outbreaks not affected by OCV campaigns
```

## Vaccine impact compuatation

### Dynamic model computation
Indirect vaccine effectiveness is a function of effective vaccine coverage
and was estimated by running dynamic models that were fitted to 
outbreaks whose time series show typical outbreak.


Additional data sets need to compute the vaccine impact
```{r}
params <- as.data.frame(data.table::fread("outputs/params_20240529.csv"))
ive_data <- as.data.frame(data.table::fread("outputs/ive_pred_20240529.csv"))
age_dist <- as.data.frame(fread("data/wpp_africa_ageprop.csv"))
```

## Vaccine impact simulation

#### Test run

For the static model, we only need to run the vaccination week 1. The impact by the vaccination on Week 2 will be just one week delayed.

```{r, eval=FALSE}
vacc_cov <- seq(0.1, 0.9, by=0.1)
vacc_week <- 1:16

test_run <- run_vacc_impact(
  outbreak_data = ts824_wk_cumsum,
  outbreak_data2 = ds,
  ive_data = ive_data,
  parameters = params,
  age_dist = age_dist,
  vacc_week = vacc_week,
  vacc_cov = vacc_cov,
  runid = 1)

(nrow(test_run))
# (nrow(test_run) == length(vacc_cov) * nrow(ts824_wk_cumsum))
(nrow(test_run) == length(vacc_week) * length(vacc_cov) * length(unique(ts824_wk_cumsum$id)))

test_run |> group_by(week_vaccination) |> reframe(ve_wk = week_delay_to_vacc_eff[1])
```

#### Sequential runs
```{r, eval=FALSE}
nruns <- 30 # number of simulation runs
# UN data on country- and year-specific age distribution (< 5 yo and 5+ yo)
age_dist <- as.data.frame(fread("data/wpp_africa_ageprop.csv"))
vacc_cov <- seq(0.1, 0.9, by=0.1)
lst_sim <- vector("list", nruns)

for (runid in 1:nruns) {

  lst_sim[[runid]] <- 
    run_vacc_impact(
      outbreak_data = ts824_wk_cumsum,
      outbreak_data2 = ds,
      ive_data = ive_data,
      parameters = params,
      age_dist = age_dist,
      vacc_week = 1,
      vacc_cov = vacc_cov,
      runid = 1)
}
# fwrite(static_vim, paste0("outputs/static_vim", tstamp(), ".csv"))
# saveRDS(lst_sim, paste0("outputs/static_vim_lst_", tstamp(), ".rds"))
```

#### Parallel `run_vacc_impact`
```{r, eval=FALSE}
tic <- Sys.time()

library(doParallel)
# UN data on country- and year-specific age distribution (< 5 yo and 5+ yo)
age_dist <- as.data.frame(fread("data/wpp_africa_ageprop.csv"))
vacc_cov <- seq(0.1, 0.9, by=0.1)
vacc_week <- 1:16
nruns <- 200 # number of simulation runs

cl <- makeCluster(getOption("cl.cores", detectCores() - 1))
registerDoParallel(cl)

svim_list <- 
  foreach(i = 1:nruns, .packages = c("data.table")) %dopar% {
    run_vacc_impact(
      outbreak_data = ts824_wk_cumsum,
      outbreak_data2 = ds,
      ive_data = ive_data,
      parameters = params,
      age_dist = age_dist,
      vacc_week = vacc_week,
      vacc_cov = vacc_cov,
      runid = i)
}
stopCluster(cl)

Sys.time() - tic
# Time difference of 36.02764 mins
saveRDS(svim_list, paste0("outputs/svim_list_", tstamp(), ".rds"))
```

#### Parallel `run_vacc_impact_weekly`
```{r, eval=FALSE}
tic <- Sys.time()

library(doParallel)
# UN data on country- and year-specific age distribution (< 5 yo and 5+ yo)
age_dist <- as.data.frame(fread("data/wpp_africa_ageprop.csv"))
vacc_cov <- seq(0.1, 0.9, by=0.1)
vacc_week <- 1:16
nruns <- 200 # number of simulation runs

cl <- makeCluster(getOption("cl.cores", detectCores() - 1))
registerDoParallel(cl)

svim_list <- 
  foreach(i = 1:nruns, .packages = c("data.table")) %dopar% {
    run_vacc_impact_weekly(
      outbreak_data = ts824_wk_cumsum,
      outbreak_data2 = ds,
      ive_data = ive_data,
      parameters = params,
      age_dist = age_dist,
      vacc_week = vacc_week,
      vacc_cov = vacc_cov,
      runid = i)
}
stopCluster(cl)

Sys.time() - tic
# Time difference of 36.02764 mins
saveRDS(svim_list, paste0("outputs/svim_week_list_", tstamp(), ".rds"))
```
