---
title: "Table 1. Epidemiologic features and transmission classification of the modeled typhoid outbreaks"
format:
  docx:
    toc: false
    reference-doc: _landscape_ref.docx
execute:
  echo: false
  warning: false
---

```{r}
d <- read.csv("outbreak_epi_summary.csv", check.names = FALSE, stringsAsFactors = FALSE)
sym <- function(x) c("–", "◐", "●")[x + 1]   # 0 none / 1 contributory / 2 primary
t1 <- data.frame(
  `Outbreak` = d$outbreak,
  `Location, country` = paste0(d$location, ", ", d$country),
  `Period` = d$period,
  `Suspected` = d$suspected,
  `BC confirmed` = d$bc_confirmed,
  `Vehicle / source` = d$vehicle,
  `Point-source` = sym(d$pt_source),
  `Environmental (water)` = sym(d$env_water),
  `Person-to-person` = sym(d$p2p),
  `theta (mode)` = sprintf("%.2f (%s)", d$theta_cs, d$mode),
  `Resolution` = d$resolution,
  `Model` = d$model,
  `Independence group` = d$indep_group,
  check.names = FALSE)
knitr::kable(t1)
```

**Route indicators:** ● primary/strong · ◐ contributory · – not evidenced.
**θ** = inferred common-source fraction (source-decomposition); 1 − θ = propagated (person-to-person).
**Model:** resolution-eligible outbreaks use one additive renewal-with-source recursion;
θ allocates baseline true typhoid between source and propagated components rather than weighting
completed static and renewal outcomes. Monthly/fortnightly resolution → static only (renewal
$R_t$ is unidentifiable below μ_g/Δ ≈ 2).
**Independence:** 19 reported series correspond to **~12 independent epidemics**; the Kasese-Uganda
(Neil + Walters ×2), Harare 2011–12 (Polonsky ×2 + Muti + Imanishi ×2) and Hyderabad-XDR
(Qamar + Yousafzai) groups each collapse multiple reports of one epidemic, and Muyembe-Tamfum 2009
is excluded from incidence modeling as a delayed intestinal-perforation (complications) series.

\newpage

## Table 1b. Classifying evidence (per outbreak)

```{r}
t2 <- data.frame(Outbreak = d$outbreak, `Classifying evidence` = d$evidence, check.names = FALSE)
knitr::kable(t2)
```
