---
title: "Typhoid ORI manuscript revision — results manifest"
subtitle: "One row per pending margin comment (c0-c9), with supporting tables and figures"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    df-print: kable
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr); library(dplyr)
root <- Sys.getenv("RENEWAL_ROOT", normalizePath(".."))
tab <- function(n) read.csv(file.path(root, "revision/tables", n), check.names = FALSE)
out <- function(n) read.csv(file.path(root, "revision/outputs", n), check.names = FALSE)
fig <- function(n) file.path(root, "revision/figures", paste0(n, ".png"))
```

## Revision manifest (primary handoff)

Each row maps a pending manuscript value to its computed result. Fill the
manuscript placeholders directly from this table.

```{r manifest}
kable(out("revision_values.csv"))
```

## Part A — Data reconciliations (c0, c5, c6)

```{r partA}
kable(tab("tab_reconciliations.csv"))
```

## Part B — Corrected global extrapolation (c7)

The submitted **1,409 / 1,461 = 96%** is impossible under the static model
(maximum proportional reduction P_max ≈ 72%). Corrected aggregate annual cases
averted at 1- and 8-week delay, 80% coverage:

```{r partB}
kable(tab("tab_global_extrapolation.csv"))
```

## Part C — Cost-effectiveness decomposition (c8)

Deaths averted are near zero, so averted DALYs are almost entirely YLD. Cost per
DALY averted sits well above 0.5×GDP (Ochalek opportunity cost), 1×GDP and 3×GDP.

```{r partC}
kable(tab("tab_ce_decomposition.csv"), digits = 2)
```

```{r partC_fig, fig.cap="Cost per DALY averted vs GDP-based thresholds."}
knitr::include_graphics(fig("fig_ce_thresholds"))
```

### c2 — Coverage-dependence of indirect VE

`IVE(f) = (OVE − f·TVE)/(1−f)` (Khanam Vi-TT: OVE=0.57, TVE=0.85) is
coverage-dependent and turns negative above f ≈ 67%. The fixed IVE = 17% is best
stated as a conservative assumption; the renewal model removes the need for it.

```{r c2}
kable(tab("tab_ive_coverage.csv"))
```

```{r c2_fig, fig.cap="Algebraic IVE(f) vs the submitted fixed 17%."}
knitr::include_graphics(fig("fig_ive_coverage"))
```

## Part D — Confirmation adjustment for suspected cases (c3)

Per-outbreak α = (confirmed/suspected) / s_culture (Mogasale 2016 = 0.61),
capped at 1. Vaccination cost is fixed; only true-case-linked quantities scale.
**Percent case reduction is invariant** to α (validated).

```{r partD}
kable(tab("tab_confirmation_adjusted.csv"), digits = 2)
```

## Part E — Spatial / hotspot targeting (c8; ILLUSTRATIVE)

Parameterized illustrative scenario (not a fitted estimate): vaccinating the
sub-area holding φ_cases = 80% of cases needs only φ_pop of the population.

```{r partE}
kable(tab("tab_spatial_targeting.csv"), digits = 1)
```

```{r partE_fig, fig.cap="Whole-unit vs targeted cost per DALY (illustrative)."}
knitr::include_graphics(fig("fig_spatial_targeting"))
```

## Part F — Statistical reporting (c9)

Ranges labelled "95% CI" are percentile/IQR spreads, not confidence intervals.
Both IQR and 2.5–97.5 percentile bounds are reported; pooled estimates include
population- and cases-weighted variants.

```{r partF}
kable(tab("tab_intervals.csv"))
kable(tab("tab_weighted_impact.csv"), digits = 1)
```

```{r partF_fig, fig.cap="Per-outbreak % reduction, static vs renewal, with pooled and population-weighted."}
knitr::include_graphics(fig("fig_forest_pctreduction"))
```

## Part G — Renewal-model results (c1)

Manuscript-ready renewal outputs (R_t panels, self-consistency/resolution tables,
forest, amplification, η_eff, policy grid, GI/ψ_T sensitivity) are produced by the
companion module `renewal/run_analysis.R` (see `renewal/report.html`).

## Key assumptions

- **GI:** gamma, mean 14 d (CV ≈ 0.6), first weekly bin dropped; renewal only where μ_g/Δ ≥ 2.
- **ψ = 0.83** (matches static pipeline); ψ_T = ψ base.
- **s_culture = 0.61** (95% CI 0.52–0.70), Mogasale et al. 2016.
- **Thresholds:** 0.5×GDP (Ochalek et al. 2018 opportunity cost), 1×GDP, 3×GDP.
- **Spatial φ:** φ_cases = 0.80, φ_pop swept 0.25–0.50 — **illustrative only**.
- **Inflation:** Sub-Saharan Africa CPI proxy (provided IMF file lacks a World row).
```
