---
title: "Case definitions as observation operators"
subtitle: "Typing each outbreak's ascertainment and designing definition-appropriate fitting"
date: today
format:
  html:
    toc: true
    embed-resources: true
    df-print: kable
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr); library(dplyr)
root <- Sys.getenv("RENEWAL_ROOT", normalizePath(".."))
rd <- function(n) read.csv(file.path(root, "case_definitions", n), check.names = FALSE)
fig <- function(n) file.path(root, "case_definitions/figures", paste0(n, ".png"))
D <- rd("tab_case_definitions.csv"); Fd <- rd("tab_fitting_design.csv")
```

## Why this matters

One latent infection process λ_o(t) per outbreak; each study observes a different transform
`observed = rho_o · (λ * d_o)`. The case definition sets the ascertainment scale rho_o and
delay d_o. Mishandling it biases absolute/cost outcomes (rho_o), the percent-reduction
invariance (needs rho_o time-constant), θ/renewal (d_o distorts curve shape), and DALYs
(severity mix). The deliverable decides — per outbreak — pool / harmonize / stratify / exclude.

## Types of the 13 renewal outbreaks

```{r typefig, fig.cap="Observation-process type of the 13 renewal outbreaks."}
knitr::include_graphics(fig("fig_type_distribution"))
```

```{r typetab}
kable(D |> filter(renewal_set == "Y") |>
  select(study_id, type, confirmation_method, definition_changed_midoutbreak,
         severity_restriction, confidence), caption = "Case definitions (renewal outbreaks).")
```

## The worked problem: Muyembe vs Neil (and the real second perforation series)

- **Muyembe-Tamfum 2009 — Type D, a peritonitis/surgical case series** (144 perforation
  admissions; the severe ~1–3% subset; long severity-filtered delay). It is in the
  renewal-13, so its R_t/θ are built on a *perforation* curve. **Recommendation: exclude
  from renewal/θ/%-reduction (sensitivity case); keep only for a perforation-mortality (YLL)
  subgroup with Walters-Kasese.**
- **Neil 2012 — Type A surveillance, NOT a perforation series** (perforation is one optional
  qualifier; IP rate 43% passive but 5–8% under enhanced surveillance). This corrects the
  prompt's premise. But **ascertainment changed mid-outbreak** → invariance violated, early
  curve IP-distorted → **include with caution**, flag the change.
- **The actual second perforation-biased series is Walters-Kasese** (de facto IP-restricted
  retrospectively, 82% IP). A perforation subgroup = **Muyembe + Walters-Kasese**.

## Fitting design (per outbreak)

```{r design}
kable(Fd |> filter(renewal_set == "Y") |>
  select(study_id, type, rho_scale, invariance_status, pool_pct_reduction, pool_theta, recommendation),
  caption = "Observation operator, invariance, poolability, recommendation (renewal outbreaks).")
```

## Poolability counts (percent-reduction / θ pool)

```{r counts}
fr <- Fd |> filter(renewal_set == "Y")
data.frame(
  pool = c("Type A pool-as-reference", "Type C suspected+rho", "Type B harmonize-absolute",
           "Include w/ caution (invariance)", "Exclude (Type D perforation)", "Flag (misfiled)"),
  n = c(sum(grepl("pool as reference", fr$recommendation)),
        sum(grepl("suspected for dynamics", fr$recommendation)),
        sum(grepl("true-case", fr$recommendation)),
        sum(grepl("caution", fr$recommendation)),
        sum(grepl("EXCLUDE", fr$recommendation)),
        sum(grepl("human review|misfiled", fr$recommendation)))) |> kable()
```

## Pipeline linkages

- **Confirmation adjustment (α):** Types A/C use α (suspected→true). **Type B (Qamar,
  Yousafzai) need the *inverse* scaling** (already confirmed). Type D needs the
  perforation-rate scale, not α.
- **θ / transmission-mode (`source_decomposition/`, `hierarchical_theta/`):** drop Muyembe;
  caution on Neil.
- **DALY/cost:** Type D carries real mortality (YLL); A/C/B morbidity-dominated.

## Honesty notes
Definitions paraphrased and cited; "not clearly stated" for Hendriksen (epi basis undefined)
(epi basis undefined); Davis read from the real MMWR (the Zotero copy was misfiled). No definition imputed. Confidence flagged per row.
```
