---
title: "Additive source and propagated-transmission renewal model"
subtitle: "Two additive levels, with the former theta-weighted calculation retained as a structural comparator"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr)
library(dplyr)
root <- Sys.getenv("RENEWAL_ROOT", normalizePath(".."))
tab <- function(n) read.csv(
  file.path(root, "source_decomposition/tables", n),
  check.names = FALSE)
fig <- function(n) file.path(
  root, "source_decomposition/figures", paste0(n, ".png"))
mode <- tab("tab_transmission_mode.csv")
cmp <- tab("tab_additive_vs_theta_weighted.csv")
pooled <- tab("tab_additive_vs_theta_weighted_pooled.csv")
validation <- tab("tab_additive_validation.csv")
fmt <- function(x, d = 1) formatC(x, digits = d, format = "f")
```

## Summary

The primary analysis uses **additive models at two different levels**.

1. **Observation level:** observed suspected febrile incidence is the sum of
   latent true typhoid and other febrile illness,

   $$
   S_t = T_t + F_t.
   $$

   A posterior PPV draw fixes the cumulative true-typhoid burden,
   $\sum_tT_t=\pi\sum_tS_t$. The non-typhoid component $F_t=S_t-T_t$ is held
   fixed under typhoid vaccination.

2. **Transmission level:** true-typhoid incidence is the sum of exogenous or
   background-source incidence and propagated incidence,

   $$
   T_t = X_t + P_t,\qquad
   P_t = R_t^P\sum_{s\ge1}w_sT_{t-s}.
   $$

   The source fraction $\theta$ partitions baseline renewal-identifiable
   incidence into $X_t=\theta T_t$ and $P_t=(1-\theta)T_t$. Positive incidence
   with zero prior infectiousness is necessarily an exogenous seed and is
   assigned to $X_t$ regardless of $\theta$.

This formulation places the two transmission components **inside one renewal
recursion**. It replaces the former calculation that averaged two completed
outcomes,

$$
A_{\mathrm{weighted}}
=\theta A_{\mathrm{static}}+(1-\theta)A_{\mathrm{renewal}}.
$$

That historical calculation remains below only as a paired structural
comparator.

## Vaccine counterfactual

Let $q_t=1-c(t)\kappa\psi_T$, where $\kappa$ is coverage, $\psi_T$ is
transmission-relevant vaccine efficacy, and $c(t)$ is the protected fraction.
The additive counterfactual is

$$
\begin{aligned}
X_t^v &= q_tX_t,\\
P_t^v &= q_tR_t^P\sum_{s\ge1}w_sT_{t-s}^v,\\
T_t^v &= X_t^v+P_t^v,\\
S_t^v &= T_t^v+F_t.
\end{aligned}
$$

Direct protection therefore applies to typhoid acquisition from both source
and propagated routes. Indirect protection emerges only through the modified
history in the propagated term. All true cases, including exogenous seeds,
contribute to subsequent infectiousness.

## Transmission-mode evidence and identifiability

The paper-derived classification remains primary because a single incidence
curve cannot reliably distinguish a source pulse from propagated transmission.
The observed-curve diagnostic is retained only as a suggestive cross-check and
may widen, but does not move, the paper-derived point value of $\theta$.

```{r modefig, fig.cap="Paper-based transmission mode of the renewal-eligible outbreaks."}
knitr::include_graphics(fig("fig_mode_distribution"))
```

```{r modetab}
kable(
  mode |>
    select(
      study_id, mode_paper, theta_paper_lo, theta_paper_hi,
      theta_data, agreement, final_theta, confidence, citation),
  digits = 2,
  caption = paste(
    "Paper evidence and observed-curve diagnostic.",
    "The paper-derived midpoint is the primary source fraction."))
```

```{r xval, fig.cap="Paper-derived versus observed-curve source-fraction diagnostics."}
if (file.exists(fig("fig_theta_crossvalidate")))
  knitr::include_graphics(fig("fig_theta_crossvalidate"))
```

## Simulation comparison with theta weighting

The additive and weighted calculations were evaluated on identical paired
draws: the same PPV, vaccine efficacy, campaign timing, coverage, generation
interval, and source fraction were used in both calculations. The exact
historical comparator retains the static model's fixed indirect-effectiveness
term $\eta=0.17$. A second, mechanism-matched comparator sets $\eta=0$ so that
its endpoints agree with the additive model: $\theta=0$ is pure renewal and
$\theta=1$ is direct-only static impact.

Across the `r pooled$n_outbreaks` renewal-eligible outbreaks, the aggregate
true-typhoid reduction was `r fmt(pooled$additive_true_pct_median)`% under the
additive recursion and
`r fmt(pooled$historical_weighted_true_pct_median)`% under the historical
theta-weighted calculation. The paired difference was
`r sprintf("%+.1f", pooled$delta_true_pp_median)` percentage points
(`r fmt(pooled$delta_true_pp_p025)` to
`r fmt(pooled$delta_true_pp_p975)` across paired uncertainty draws). Aggregate
true cases averted were `r fmt(pooled$additive_cases_median, 0)` versus
`r fmt(pooled$historical_weighted_cases_median, 0)`.

```{r comparison-figure, fig.cap="Primary additive recursion versus the historical theta-weighted calculation. Blue horizontal bars show the structural range obtained from each outbreak's source-fraction bounds; they are not credible intervals."}
knitr::include_graphics(fig("fig_additive_vs_theta_weighted"))
```

```{r difference-figure, fig.cap="Paired percentage-point difference between the additive and historical weighted formulations."}
knitr::include_graphics(fig("fig_additive_difference"))
```

```{r comparison-table}
cmp_show <- cmp |>
  transmute(
    Outbreak = study_id,
    `theta` = theta,
    `PPV` = ppv_median,
    `Additive %` = pct_additive_median,
    `Historical weighted %` = pct_theta_weighted_historical_median,
    `Mechanism-matched weighted %` = pct_theta_weighted_matched_median,
    `Paired difference, percentage points` =
      delta_true_pp_vs_historical_median,
    `Observed suspected % (additive)` =
      observed_pct_additive_median)
kable(
  cmp_show, digits = 1,
  caption = paste(
    "Per-outbreak medians at campaign week 8 and 80% coverage.",
    "Zeros indicate that effective protection occurred after the modeled",
    "outbreak tail under the prespecified timing rule."))
```

The aggregate difference is small, but that does not imply equivalence for
individual outbreaks. The additive recursion allows source-generated cases to
seed later propagated cases and lets vaccination interrupt that coupled
process. The weighted calculation instead treats the source and propagated
outcomes as independent completed branches. The difference can therefore be
larger, and can change sign, for particular epidemic shapes and response times.

## Validation

```{r validation}
kable(validation, digits = 12)
```

All checks must pass before the driver writes a completed result. They cover:

- $S_t=T_t+F_t$ and cumulative PPV conservation;
- $T_t=X_t+P_t$ at baseline and under vaccination;
- non-negative components and no vaccine-induced increase in incidence;
- the $\theta=1$ direct-only endpoint;
- exact reproduction of both frozen weighted formulas; and
- uniqueness of all paired simulation keys.

The source-fraction range is a **structural sensitivity interval**, not a
posterior credible interval. PPV and vaccine-parameter intervals describe
posterior or sampled uncertainty conditional on that structural choice.

## Limitations

- The temporal allocation of true typhoid is derived from a flat-background
  candidate followed by non-negative truncation and PPV-conserving
  renormalisation. The resulting residual $F_t$ is fixed under vaccination but
  is not generally perfectly flat.
- A constant outbreak-level $\theta$ partitions renewal-identifiable incidence.
  Time-varying contamination or route-specific vaccine effects would require
  additional data.
- Source fractions are based primarily on field-investigation evidence. The
  hierarchical renewal-with-source analysis failed its prespecified
  identifiability gate, so curve-derived posterior source fractions are not
  substituted here.
- Stream B is computed from the observed suspected curve and is only a
  diagnostic; it is not a fitted decomposition of latent true-typhoid incidence.
