---
title: "Transmission-mode classification & source decomposition"
subtitle: "Common-source (static) vs propagated (renewal) vs mixed — and the static↔renewal bracket"
date: today
format:
  html:
    toc: true
    embed-resources: true
    df-print: kable
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr); library(dplyr)
root <- Sys.getenv("RENEWAL_ROOT", normalizePath(".."))
tab <- function(n) read.csv(file.path(root, "source_decomposition/tables", n), check.names = FALSE)
fig <- function(n) file.path(root, "source_decomposition/figures", paste0(n, ".png"))
```

## Question

The static model is correct for **common-source** outbreaks (a water/food source drives
incidence; transmission is non-reproductive) and the renewal model for **propagated**
(person-to-person) outbreaks. We classify each of the 13 weekly-resolvable outbreaks,
assign a source fraction θ (1 = pure common-source → static; 0 = pure propagated →
renewal), and report impact as the **static↔renewal bracket** with a θ point estimate.

Two evidence streams are triangulated: **A — paper evidence** (full-text reading of each
primary report against a fixed rubric) and **B — curve-shape diagnostics** computed from the
incidence series (common-source fit + propagation signals).

## Headline

```{r modefig, fig.cap="Paper-based transmission mode of the 13 renewal outbreaks."}
knitr::include_graphics(fig("fig_mode_distribution"))
```

Reading the actual reports **revises the optimistic seeded demonstration**: most of these
outbreaks are **mixed** (an identified water source *and* documented onward transmission /
resurgence), several are explicitly described by their authors as "propagated" (Muti,
Qamar, Yousafzai), and one is clearly propagated (Neil 2012, Kasese — no source implicated,
~19-month multi-strain epidemic). Only a few are clean common-source (Lewis 2005, with
S. Typhi isolated and PFGE-matched to the tap; Michel 2005, a food point-source [excluded
from the pipeline]). **Implication: the dataset is *not* source-dominated — so the
static↔renewal bracket (not a pure-static claim) is the defensible framing.**

## Classification & cross-validation

```{r modetab}
kable(tab("tab_transmission_mode.csv") |>
        dplyr::select(study_id, mode_paper, theta_paper_lo, theta_paper_hi,
                      theta_data, agreement, final_theta, confidence, citation),
      digits = 2, caption = "Stream A (paper) vs Stream B (data) source fraction theta.")
```

```{r xval, fig.cap="Cross-validation of theta: paper (Stream A) vs data (Stream B)."}
if (file.exists(fig("fig_theta_crossvalidate"))) knitr::include_graphics(fig("fig_theta_crossvalidate"))
```

**Identifiability caveat.** A single incidence curve cannot definitively separate a
common-source pulse (convolved with the incubation period) from propagation — they can
produce identical curves (the same non-identifiability the manuscript notes). Stream B is
therefore *suggestive*, and the paper evidence (Stream A) is the primary classifier.

## The static↔renewal bracket

```{r bracket, fig.cap="Per-outbreak bracket: renewal (theta=0) to static (theta=1) with the theta point."}
knitr::include_graphics(fig("fig_decomposition_bracket"))
```

```{r dectab}
kable(tab("tab_decomposition.csv") |>
        dplyr::select(study_id, final_theta, pct_renewal, pct_static, pct_theta),
      digits = 1, caption = "% cases averted: renewal endpoint, static endpoint, and theta point (tau=8, 80%).")
```

The decomposition satisfies `averted(theta) = theta·static + (1-theta)·renewal`, validated so
that θ=1 reproduces the static model and θ=0 the renewal model to machine precision.

## Limitations

- Stream A depends on report completeness; several reports are short (MMWR notes) or are
  genomic/clinical papers thin on transmission mode (Hendriksen, Muyembe) → low confidence.
- **Davis 2018** is read from the real MMWR (the Zotero copy was misfiled): waterborne
  common-source (boreholes 13/18 fecal-contaminated; cases clustered around boreholes;
  declined after WASH) → mode A.
- The protocol's outbreak list (Supp. Table 1) and the analysis dataset differ slightly
  (it lists Michel 2005, which the pipeline excludes, and omits Lewis 2005, which the
  pipeline includes); both are reconciled here.
- Stream B's θ is heuristic given the identifiability limit; θ is treated as a bracket with
  sensitivity, not a point parameter.
```
