---
title: "Renewal-equation model of typhoid outbreak-response immunization"
subtitle: "Additive observation and transmission-source formulations"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-fold: true
    fig-width: 9
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr)
library(dplyr)
root <- Sys.getenv("RENEWAL_ROOT", normalizePath(".."))
tab <- function(n) read.csv(
  file.path(root, "renewal/tables", n), check.names = FALSE)
fig <- function(n) file.path(
  root, "renewal/figures", paste0(n, ".png"))
src_tab <- function(n) read.csv(
  file.path(root, "source_decomposition/tables", n),
  check.names = FALSE)
src_fig <- function(n) file.path(
  root, "source_decomposition/figures", paste0(n, ".png"))
res <- readRDS(file.path(root, "renewal/outputs/results.rds"))
have_source_results <- file.exists(
  file.path(
    root, "source_decomposition/tables",
    "tab_additive_vs_theta_weighted_pooled.csv"))
if (have_source_results) {
  source_pooled <- src_tab("tab_additive_vs_theta_weighted_pooled.csv")
  source_cmp <- src_tab("tab_additive_vs_theta_weighted.csv")
  source_validation <- src_tab("tab_additive_validation.csv")
}
fmt <- function(x, d = 1) formatC(x, digits = d, format = "f")
```

## Model overview

The analysis uses additive models at two distinct levels.

### Level 1: true typhoid versus other febrile illness

Observed suspected incidence is

$$
S_t=T_t+F_t,
$$

where $T_t$ is latent true-typhoid incidence and $F_t$ is other febrile illness
meeting the suspected-case definition. For each posterior PPV draw,
$\sum_tT_t=\pi\sum_tS_t$. The residual $F_t=S_t-T_t$ is held fixed under
vaccination, so the vaccine cannot avert non-typhoid fever.

### Level 2: source versus propagated typhoid transmission

True-typhoid incidence is

$$
T_t=X_t+P_t,\qquad
P_t=R_t^P\Lambda_t,\qquad
\Lambda_t=\sum_{s\ge1}w_sT_{t-s},
$$

where $X_t$ is exogenous or background-source typhoid and $P_t$ is propagated
typhoid. If $\theta$ is the source fraction, it partitions baseline
renewal-identifiable incidence inside this equation. It is **not** used to
average completed static and renewal outcomes.

With $q_t=1-c(t)\kappa\psi_T$, the vaccination counterfactual is

$$
X_t^v=q_tX_t,\qquad
P_t^v=q_tR_t^P\sum_{s\ge1}w_sT_{t-s}^v,\qquad
T_t^v=X_t^v+P_t^v.
$$

Direct protection acts on typhoid acquisition from both routes. Indirect
protection emerges through the propagated term's modified history. Positive
incidence with no prior infectiousness is treated as an exogenous seed.

## Renewal eligibility and baseline reproduction

The renewal model is used only when the generation interval spans at least two
reporting intervals. Monthly series are therefore excluded from the dynamic
analysis.

```{r resolution}
kable(tab("tab_resolution.csv"))
```

With vaccine efficacy set to zero, forward propagation must reproduce the
observed baseline.

```{r selfconsistency}
kable(tab("tab_selfconsistency.csv"), digits = 16)
```

## Pure-renewal and static reference results

The standard renewal run remains useful as the $\theta=0$ propagated endpoint
and for timing, coverage, and generation-interval sensitivities. The static
model remains a reference calculation. It is not mixed with renewal outputs to
form the primary source-decomposed estimate.

```{r forest, fig.cap="Per-outbreak percentage reduction under the static and pure-renewal reference calculations."}
knitr::include_graphics(fig("fig_forest_pctreduction"))
```

```{r pooled}
kable(
  res$pooled |> filter(tau == 8),
  digits = 1,
  caption = "Static and pure-renewal reference summaries at the base timing.")
```

```{r rt, fig.cap="Exact reconstruction and EpiEstim cross-check."}
knitr::include_graphics(fig("fig_Rt_panels"))
```

```{r timing, fig.cap="Pure-renewal versus static cases averted across response timing."}
knitr::include_graphics(fig("fig_amplification_vs_delay"))
```

## Additive transmission result versus theta weighting

```{r source-result}
if (!have_source_results) {
  cat(paste(
    "Source-comparison outputs are absent. Run",
    "`Rscript source_decomposition/run_source_decomposition.R`,",
    "then render this report again."))
} else {
  cat(paste0(
    "Across ", source_pooled$n_outbreaks,
    " renewal-eligible outbreaks, the aggregate true-typhoid reduction was ",
    fmt(source_pooled$additive_true_pct_median), "% under the additive recursion and ",
    fmt(source_pooled$historical_weighted_true_pct_median),
    "% under the historical theta-weighted calculation. The paired difference was ",
    sprintf("%+.1f", source_pooled$delta_true_pp_median),
    " percentage points (",
    fmt(source_pooled$delta_true_pp_p025), " to ",
    fmt(source_pooled$delta_true_pp_p975), ")."))
}
```

The historical comparator is frozen as

$$
A_{\mathrm{weighted}}
=\theta A_{\mathrm{static}}+(1-\theta)A_{\mathrm{renewal}},
$$

including its original fixed static indirect-effectiveness term
$\eta=0.17$. A second mechanism-matched comparator uses $\eta=0$ to isolate
the difference between adding components inside the recursion and averaging
completed outcomes.

```{r source-figure, fig.cap="Primary additive source-plus-propagated recursion versus the historical theta-weighted outcome calculation."}
if (have_source_results)
  knitr::include_graphics(src_fig("fig_additive_vs_theta_weighted"))
```

```{r source-difference, fig.cap="Paired percentage-point difference between the additive and historical weighted formulations."}
if (have_source_results)
  knitr::include_graphics(src_fig("fig_additive_difference"))
```

```{r source-table}
if (have_source_results) {
  kable(
    source_cmp |>
      transmute(
        Outbreak = study_id,
        theta = theta,
        `Additive %` = pct_additive_median,
        `Historical weighted %` =
          pct_theta_weighted_historical_median,
        `Difference, percentage points` =
          delta_true_pp_vs_historical_median,
        `Observed suspected % (additive)` =
          observed_pct_additive_median),
    digits = 1)
}
```

The pooled difference is smaller than the differences for some individual
outbreaks. The additive recursion permits source-generated cases to seed later
propagated cases; the weighted calculation instead treats the two outcomes as
independent branches. Differences can therefore vary in magnitude and sign
with epidemic shape and intervention timing.

## Sensitivity analyses

```{r gi, fig.cap="Generation-interval sensitivity for the pure-renewal endpoint."}
knitr::include_graphics(fig("fig_gi_sensitivity"))
```

```{r psit, fig.cap="Transmission-efficacy sensitivity for the pure-renewal endpoint."}
knitr::include_graphics(fig("fig_psiT_sensitivity"))
```

## Validation and interpretation

```{r source-validation}
if (have_source_results)
  kable(source_validation, digits = 12)
```

The source-fraction range is a structural sensitivity range, not a posterior
credible interval. PPV and vaccine-parameter intervals are conditional on that
structural assignment. Curve shape alone did not identify source fractions in
the prespecified hierarchical gate, so paper evidence remains primary.

The Level-1 allocation begins with excess above a flat other-febrile candidate,
then applies non-negative truncation and PPV-conserving renormalisation.
Therefore the final fixed residual $F_t$ is not generally perfectly flat.
