---
title: "Renewal-equation (R_t) model of typhoid ORI: results"
subtitle: "Dynamic complement to the static proportional-reduction model"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-fold: true
    fig-width: 9
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr); library(dplyr)
root <- Sys.getenv("RENEWAL_ROOT", normalizePath(".."))
tab <- function(n) read.csv(file.path(root, "renewal/tables", n), check.names = FALSE)
fig <- function(n) file.path(root, "renewal/figures", paste0(n, ".png"))
res <- readRDS(file.path(root, "renewal/outputs/results.rds"))
```

## Summary

This report compares two models of typhoid outbreak-response immunization (ORI)
impact on an identical set of `r length(unique(res$summ_grid$study_id))`
daily/weekly outbreaks and identical Sobol parameter draws:

- **Static** (manuscript Eq. 2): a fixed proportional reduction with an exogenous
  indirect-effectiveness term η = 17 %.
- **Renewal** (this work): the vaccine reduces transmission (R_t), and herd
  protection **emerges endogenously** from forward propagation — no η is supplied.

The only quantity that differs between the models is how cases-averted is computed;
DALY and cost calculations are shared (reused from `R/2-functions.R`).

## 1. Which outbreaks qualify (resolution rule)

The renewal model is identifiable only where the generation interval spans several
reporting intervals (μ_g/Δ ≳ 2). Monthly series collapse ~95 % of GI mass into one
reporting bin and are excluded; daily/weekly series qualify.

```{r resolution}
kable(tab("tab_resolution.csv"))
```

## 2. Self-consistency

With ψ = 0 the renewal counterfactual must reproduce the observed curve exactly
(the observed outbreak *is* the baseline).

```{r selfconsistency}
kable(tab("tab_selfconsistency.csv"), digits = 16)
```

## 3. Base case (τ = 8 wk, π = 0.80): static vs renewal

```{r forest, fig.cap="Per-outbreak % cases averted, static vs renewal."}
knitr::include_graphics(fig("fig_forest_pctreduction"))
```

```{r pooled}
kable(res$pooled %>% filter(tau == 8), digits = 1,
      caption = "Pooled and population-weighted % reduction at base case.")
```

## 4. Reconstructed R_t

```{r rt, fig.cap="Exact reconstruction (black) vs EpiEstim 95% CrI (blue)."}
knitr::include_graphics(fig("fig_Rt_panels"))
```

## 5. Amplification by response timing

The renewal/static ratio of cases averted exceeds 1 for early response (several
transmission generations remain) and converges to 1 for late response.

```{r amp, fig.cap="Renewal / static cases-averted ratio vs ORI delay."}
knitr::include_graphics(fig("fig_amplification_vs_delay"))
```

## 6. Effective indirect VE (η_eff)

The fixed indirect VE the static formula would need to reproduce the renewal total
reduction. η_eff > 17 % means the fixed assumption under-credits the emergent herd effect.

```{r etaeff, fig.cap="η_eff per outbreak vs the assumed 17%."}
knitr::include_graphics(fig("fig_eta_eff"))
```

```{r etatab}
kable(tab("tab_eta_eff.csv"), digits = 2)
```

## 7. Policy grid and impact summary

```{r policy, fig.cap="Impact metrics by delay × coverage, renewal vs static."}
if (file.exists(fig("fig_policy_grid"))) knitr::include_graphics(fig("fig_policy_grid"))
```

```{r impact}
kable(tab("tab_impact_summary.csv"), digits = 1,
      caption = "Impact summary (mirrors Supplementary Table 2, with renewal column).")
```

## 8. Sensitivity analyses

```{r gi, fig.cap="GI mean sensitivity (renewal % reduction)."}
knitr::include_graphics(fig("fig_gi_sensitivity"))
```

```{r psit, fig.cap="Transmission-VE (ψ_T) sensitivity (renewal % reduction)."}
knitr::include_graphics(fig("fig_psiT_sensitivity"))
```

## 9. Example counterfactual curves

```{r curves, fig.cap="Observed vs static and renewal counterfactuals."}
knitr::include_graphics(fig("fig_curves_examples"))
```

## Assumptions

- Incidence = `total_cases` (relabeled `suspected_cases`, as in `R/1-data.R`).
- GI gamma, base mean 14 d (CV ≈ 0.6), first weekly bin zeroed then renormalized.
- ψ = 0.83 (matches the static Sobol pipeline); ψ_T = ψ base, floor 0.6·ψ.
- Direct effect only in the renewal model; herd protection emerges from propagation.
```
