---
title: "Literature anchors for the transmission reproduction number R_tr"
subtitle: "Field measurements of person-to-person transmission, independent of curve shape"
date: today
format:
  html:
    toc: true
    embed-resources: true
    df-print: kable
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(knitr); library(dplyr)
root <- Sys.getenv("RENEWAL_ROOT", normalizePath(".."))
rd <- function(n) read.csv(file.path(root, "rtr_anchors", n), check.names = FALSE)
fig <- function(n) file.path(root, "rtr_anchors/figures", paste0(n, ".png"))
A <- rd("tab_Rtr_anchors.csv"); ren <- A[A$renewal_set == "Y", ]
```

## Purpose

The source fraction θ was unidentifiable from curve shape alone (`hierarchical_theta/`):
the renewal transmission term R_tr and the common-source pulse occupy the same ridge.
Field measurements of **person-to-person transmission** — household secondary attack
rates, contact-tracing counts, attributable fractions — are observed *independently* of
the incidence curve, so they can constrain R_tr along the unidentified direction.

## Coverage: the evidence is thin

```{r coverage, fig.cap="Independent R_tr anchors among the 13 renewal outbreaks."}
knitr::include_graphics(fig("fig_anchor_coverage"))
```

```{r covtab}
ren %>% count(anchor_type, name = "outbreaks") %>% kable()
```

Only **2/13** outbreaks (Polonsky's two suburbs) give a level-anchoring quantity with a
denominator; **3/13** give ordinal contact ORs (no denominator); **8/13** have nothing and
fall back to a pooled prior. **No outbreak reports a household SAR with a denominator**, so
the proposed `secondary_cases ~ Binomial(contacts, p)` likelihood term has no data to attach.

## Per-outbreak anchors

```{r anchors}
kable(A |> select(study_id, renewal_set, contact_OR, prop_secondary, anchor_type, confidence, citation))
```

## Pooled fallback (broader literature)

```{r priors}
kable(rd("tab_typhoid_transmission_priors.csv"))
```

## Constraint for the model

```{r model}
kable(rd("tab_Rtr_priors_for_model.csv"))
```

## Honest bottom line

- Independent evidence points to **low household transmission** (~5–12.5%), consistent with
  source/community-dominated typhoid.
- **Critical caveat:** these anchors measure **short-cycle household** transmission. The
  model's R_tr also includes **long-cycle community (water-mediated)** transmission — the
  dominant propagated channel here and the part confounded with the source. No field
  measurement isolates it.
- Therefore: adding these anchors to `hierarchical_theta/` and re-running the Phase-1 gate is
  worth doing, but should be expected to tighten only the **household** channel — unlikely on
  its own to make total θ identifiable. The article classification (`source_decomposition/`)
  remains primary.
```
